True versus apparent shapes of bow shocks and bow waves

### 2017-10-16 papers/Paper1/quadrics-bowshock.tex
Add section on standing waves
Write about the standing wave perturbations
Finish fig captions for shapes and projected shapes.  Finish text
about shapes.

% Standard LaTeX packages
% \usepackage[varg]{txfonts}
\usepackage[varg]{newtxmath}
\usepackage{newtxtext}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

@@

@@

\title[Bowshock shapes]{True versus apparent shapes of bow shocks and bow waves}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Tarango Yong \& Henney]{
  Jorge A. Tarango Yong \& William J. Henney\\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}
  Astrophysical bow shocks are a common result of the interaction
  between supersonic plasma flows, such as winds or jets from stars or
  active galaxies.  Similarly shaped radiation bow waves are produced
  by the relative motion between dust grains and a luminous source.
  For cylindrically symmetric bow shocks and waves, we develop a
  general theory of the effects of inclination angle, and propose a
  new two-dimensional classification scheme for bow shapes, which is
  based on dimensionless geometric ratios that can be estimated from
  observational images.  The two ratios are related to the
  ``snubness'' of the bow's apex and the ``openness'' of its wings.

@@

\input{sec-intro}
\input{sec-generic-model}
\input{sec-projection}
\input{sec-conic}
\input{sec-thin-shell}
\input{sec-conclusions}
\clearpage
\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}
\appendix
\input{app-parabola}
\input{app-rcurv}
\input{app-r90}
\input{app-wilkinoid}
\input{app-dust-wave}

>>>>

% Standard LaTeX packages
% \usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

@@

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varg]{newtxmath}

@@

\title[Bow shock shapes]{True versus apparent shapes of bow shocks and bow waves}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Tarango Yong \& Henney]{
  Jorge A. Tarango Yong \& William J. Henney\\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}
  Astrophysical bow shocks are a common result of the interaction
  between supersonic plasma flows, such as winds or jets from stars or
  active galaxies.  Similarly shaped radiation bow waves are produced
  by the relative motion between dust grains and a luminous source.
  For cylindrically symmetric bow shocks and waves, we develop a
  general theory of the effects of inclination angle, and propose a
  new two-dimensional classification scheme for bow shapes, which is
  based on dimensionless geometric ratios that can be estimated from
  observational images.  The two ratios are related to the
  ``flatness'' of the bow's apex and the ``openness'' of its wings.

@@

\input{sec-intro}
\input{sec-generic-model}
\input{sec-projection}
\input{sec-conic}
\input{sec-thin-shell}
\clearpage
\input{app-dust-wave}
\clearpage
\input{sec-standing-wave}
\input{sec-conclusions}
\clearpage
\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}
\appendix
\input{app-parabola}
\input{app-rcurv}
\input{app-r90}
\input{app-wilkinoid}

### 2017-10-16 papers/Paper1/sec-standing-wave.tex
Write about the standing wave perturbations
Finish fig captions for shapes and projected shapes.  Finish text
about shapes.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/compare_xyprime_wave-wilkinoid}
  \caption{Small-amplitude standing wave perturbations to bow shapes.}
  \label{fig:perturb-shapes}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]
  {figs/wave_xyprime-A005-N20-ancantoid-xi080-beta000500}
  \caption{Plane-of-sky projections of perturbed bow shapes}
  \label{fig:perturb-xy-prime}
\end{figure*}

@@

For simplicity, we consider standing waves with an amplitude that
multiplies the bow shock radius \(R\) and a pattern that is periodic
in the axial angle \(\theta\): \(R(\theta) \to (1 + \Delta) R(\theta)\), where
\begin{equation}
  \label{eq:standing-wave}
  \Delta = A \cos(N \theta) \cos(2\pi \varphi)
\end{equation}

>>>>

\begin{figure*}[t!]
  \centering
  \includegraphics[width=\linewidth]{figs/compare_xyprime_wave-wilkinoid}
  \caption{Small-amplitude standing wave perturbations to wilkinoid
    bow shapes.  The maximum deviations from the base shape are seen
    at phases \(\phi = 0\) (blue line) and \(\phi = 0.5\) (red line), while
    the perturbation is zero at \(\phi = 0.25\) and \(0.75\) (black
    line).  Results are shown left to right for increasing wave
    numbers \(N\) and decreasing amplitudes \(A\): (a)~\(A = 0.1\),
    \(N = 1.0\), (b)~\(A = 0.05\), \(N = 2.0\), (a)~\(A = 0.02\),
    \(N = 5.0\).}
  \label{fig:perturb-shapes}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]
  {figs/wave_xyprime-A005-N20-ancantoid-xi080-beta000500}
  \caption{Plane-of-sky projections of perturbed bow shapes.  In all
    cases, the base bow shape is ancantoid with \(\xi = 0.8\),
    \(\beta = 0.005\) and the perturbation is the same as in the central
    panel of Fig.~\ref{fig:perturb-shapes}, with amplitude
    \(A = 0.05\) and wave number \(N = 2.0\). Results are shown for
    inclination angles \(i = 0\) to \(i = 75^\circ\) (indicated by line
    color and thickness, see key) and for different fractional phases
    of the oscillation: (a)~\(\varphi = 0.0\), (b)~\(\varphi = 0.25\),
    (c)~\(\varphi = 0.50\). Unlike in Fig.~\ref{fig:perturb-shapes},
    the spatial coordinates are normalized to the instantaneous
    projected apex radius \(R_0'\) at each phase.}.
  \label{fig:perturb-xy-prime}
\end{figure*}

@@

@@

We consider fractional perturbations \(\Delta(\theta, t)\) of a base shape
\(R(\theta)\), such that
\(R(\theta) \to [1 + \Delta(\theta, t)] R(\theta)\).  For simplicity,
\(\Delta(\theta, t)\) is a standing wave of constant amplitude \(A\), which is
periodic in \(\theta\), with wave number \(N\).  In cylindrical symmetry
\(\Delta(\theta, t)\) must be even in \(\theta\), so can be expressed as
\begin{equation}
  \label{eq:standing-wave}
  \Delta(\theta, t) = A \cos(N \theta) \cos(2\pi \varphi) . 
\end{equation}
For waves with period \(P\), the fractional phase \(\varphi\) will
vary with time \(t\) as
\begin{equation}
  \label{eq:fractional-phase}
  \varphi(t) = (\varphi_0 + t/P) \bmod 1.0\ ,
\end{equation}
where \(\varphi_0\) is an arbitrary reference phase.

Example oscillations with wave numbers \(N = 1.0\), \(2.0\), and
\(5.0\) superimposed on a wilkinoid base shape are shown in
Figure~\ref{fig:perturb-shapes}.  There are \(N\) nodes of the
oscillation between \(\theta = 0 \to \pi\), always with an antinode at the apex
(\(\theta = 0\)), as required by symmetry.  So, with \(N = 1.0\) there is a
node (fixed point) in the near wing at \(\theta = \pi/2\), but an antinode in
the far wing at \(\theta = \pi\), which is in antiphase with the oscillation
of the apex, giving rise to a large-scale ``breathing'' mode of
oscillation.  With \(N = 2.0\), there are nodes at \(\theta = \pi/4\) and
\(3\pi/4\), while the antiphase antinode has moved to the near wing at
\(\theta = \pi/2\).  There is still an antinode in the far wing at
\(\theta = \pi\) but it is now in phase with the apex, giving rise to a
``curling-up/straightening-out'' mode of oscillation.  With
\(N = 5.0\), there are many more nodes and antinodes, giving a
``ringing'' mode of oscillation.  Note that all our examples have
\(A \propto 1/N\) in order to keep the local curvature relatively low.  If
the product \(A N\) is not small compared to unity, then the local
curvature can be so extreme as to reverse the concave shape of the
base bow shape, producing locally convex regions.
